---
title: "Machine Learning - Prediction Model"
author: "Henry Marroquin"
date: "Saturday, September 26, 2015"
output: html_document
---

***

###OVERVIEW
Using exercise accelerometers data provided by [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har), build a Machine Learning algorithm to predict if the subject perfomed the exercise in a correct or incorrect manner(Participants in the dataset were asked to perform barbell lifts in 5 different ways each.  ie Correctly or incorrectl).

The Training data set used for this project can be found here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The Test data set used to test the model can be found here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)


###DATA PROCESSING
Load required packaged and datasets for training a model and do subsequent out of sample testing :
```{r, cache=TRUE}
# packages:
library(caret)
library(randomForest)
library(ggplot2)


# Load training data:
rawtrain = read.csv("pml-training.csv")
# create a sub partition to do mock testing and derive an out of sample error of the model.   
intrain <- createDataPartition(y=rawtrain$classe, p=0.75, list = FALSE)
training <- rawtrain[intrain,]
# OutofSample Test:
testing <- rawtrain[-intrain,] 

```

##DATA PREPROCESSING
clean up data and remove irrelevant columns.

```{r, cache=TRUE}
#identify amount of column with NA
na_columns = sapply(training, function(x) {sum(is.na(x))})
table(na_columns)
# flag which columns to remove
remove_columns = names(na_columns[na_columns==14396])
#remove columsn
training = training[, !names(training) %in% remove_columns]
#str(training)   #used to observe new results


```

###FIT TRAINING MODEL
```{r,cache=TRUE}
modelfit <- train(classe ~.,data=training, method = "rf",prox=TRUE)

```

Look at our Model stats.  It was bootstrapped a total of 25 times, opting for a model with the highest accuracy of roughly 98%.
```{r}
modelfit
```
### TEST PREDICT MODEL WITH OUT OF SAMPLE DATA.

```{r, cache=TRUE}
#clean up testing dataset as well.
testing <- testing[, !names(testing) %in% remove_columns]
# run Prediction on out of sample dataset
predictions <- predict(modelfit,newdata=testing)
# analyse results
summary(predictions)
```

### PERFORM FULL TEST
run model on test values provided:
```{r, cache=TRUE}
rawtest = read.csv("pml-testing.csv")
#fulltest <- rawtest[, !names(rawtest) %in% remove_columns]
#Fullpredictions <- predict(modelfit,newdata=fulltest)
#summary(Fullpredictions)
```
